\begin{flushleft}
\medskip
\Large\textrm{Acknowledgments}
\medskip
\end{flushleft}
My sincere gratitude goes to Paul Lukowicz, my thesis advisor,
who is often on a tight schedule. Yet, he always makes time to provide valuable research input and inspiration.
I can always rely on his support and vast experience.

Additionally, I would like to give a warm thank you to Hans Gellersen. I'm proud to have him as second advisor, he's an extraordinary researcher and I valued his input and help throughout my Ph.D. studies.

Thanks to David Bannach, Gernot Bahle, Josef Neuburger, Gerald Pirkl, Jens Weppner and Kamil Kloch, for being great peers, for critic and proofreading, Georg Ogris, Tobias Franke, Jinyuan Cheng, Agnes Gr\"unerbl,  Michael Barry, Karl Stockinger and Peter Barth for being outstanding colleagues and for sparking interesting discussions.

Special thanks go to Ian Cloete, Christoph Schuba, Helmut Stadm\"uller and Ernst Heinz, all exceptional and inspiring teachers. Without their guidance, I would definitely not work in research today.

Thanks to Thad Starner, Albrecht Schmidt and Michael Beigl for giving valuable advice and sharing their expertise.
Other friends and inspiring colleagues, I want to thank: Oliver Amft, Ulf Blanke, Todd Farrell, Martin Kurserow, Till Riedel, Michael Springmann, Kristof Van Laerhoven and Jamie Ward.

Cyrille Thouvenin, Roland Westrelin, Eric Lemoine and Nicolas Fugier  thanks for an amazing time in Grenoble.

Kurt Partridge, Bo Begole, Rowan Nairn and Max Van Kleek thank you for a great summer at PARC.

Holger, Melanie, Sanjit and Julia deserve also a thank you for their close friendship and for putting up with me over the last couple of years.

Most importantly, I sincerely thank Renate Kunze, my mother, and Anita and Helmut Woll, my grandparents, 
for their love and support. No matter what, I can always count on them. Without them this work would not have been possible.
\newline
\newline
\newline
\newline
\newline
\noindent Kai Kunze \newline
\noindent Passau, 2011
